\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{mathtools}
\usepackage{amsmath}

\title{BAO detection in cross-correlation with DESI}
\author{Alkistis, José, Sefa, Zhaoting}
\date{May 2026}

\begin{document}

\maketitle

\section{Theoretical Model for ${\rm T \times g}$ Power Spectrum}
    To model the real space power spectrum, we use a rudimentary non-linear model.The linear power spectrum is decomposed into a sum of a smooth component, $P_{\rm nw}$ and a distinct "wiggly" part $P_{\rm w}$. This separation is done by following a "de-wiggling" routine, where two cubic splines are fitted through inflection points of the normalised power spectrum. This procedure follows the standard from BOSS\\
    \begin{equation}
        P^{\rm lin}_{\rm mm} = P_{\rm nw} + P_{\rm w}\:.
    \end{equation}
    To compute the quasi non-linear power spectrum, the wiggle power spectrum is suppressed in a $k$-dependent way. The suppression factor is a phenomenological approximation to the correct LPT result.
    \begin{equation}
        g(k, \mu) = \exp\left[-\sigma_v^2\,k^2\,\left(\left(1-\mu^2\right) + \mu^2\,\left(1 + f\right)^2\right)\right]\:,
    \end{equation}
    where $\mu$ is the angle between the line of sight and the $k$ mode in a small angle approximation. The real space variance of the velocity dispersion $\sigma_v$ can either be handled as a free parameter or fixed to the linear prediction. The growth rate $f$ is fixed to the cosmology value.\\
    The quasi non-linear matter power spectrum is given by:
    \begin{equation}
        P_{\rm mm}^{\rm QNL}(k, \mu) = P_{\rm nw}(k, \mu) + g(k, \mu)\,P_{\rm w}(k, \mu)\:.
    \end{equation}
    To obtain the tracer power spectrum in redshift space we use the linear kaiser formulation and correct for the fingers of god effect by multiplying a phenomenological function.
    \begin{equation}
        \mathrm{FOG}(k, \mu) = \frac{1}{1 + \left[f\,k\,\mu\,\sigma_p\right]^2}\:,
    \end{equation}
    where $\sigma_p$ is the real space variance of the peculiar motion of collapsed objects. We fix this quantity to the linear perturbation theory prediction. The exact form of the fingers of god suppression is not to important but the formula we use here is the same as in the state-of-the-art analysis of DESI.\\
    Since we are computing the temperature cross galaxy power spectrum we need to introduce two bias parameters.
    \begin{equation}
        P_{\rm Tg}(k, \mu) = {\rm FOG}(k, \mu) \, (b_{\rm T} + f\,\mu^2)\,(b_{\rm g} + f\,\mu^2)\,P_{\rm mm}^{\rm QNL}(k, \mu)\:.
    \end{equation}
\section{Observational Effects}
    Due to the finite survey geometry the possible $\boldsymbol{k}$ are discretized into integer multiples of the fundamental mode per dimension. The smallest accessible scale is given by the sampling rate. For a Line intensity mapping experiments this is essentially the voxel size.\\
    Additionally, the instrument Beam, the pixelation, and griding does additionally suppress the measured signal. As all three effects influence the measured temperature field while the Galaxy field is only affected by the griding, the cross power spectrum is given by
    \begin{equation}
        P_{\rm Tg}^{\rm obs\,eff}(k_i, \mu_i) = \mathcal{B}(k_i, \mu_i)\, \mathcal{P}(k_i, \mu_i)\, \mathcal{G}^2(k_i, \mu_i)\,P_{\rm Tg}(k_i, \mu_i)\:,
    \end{equation}
    where $\mathcal{B}$, $\mathcal{P}$, and $\mathcal{G}$ are the three distinct effects, respectively.\\
    Finally, the non-uniform sampling of the lightcone, leads to non-trivial survey weights which propagate to the measured power spectrum in the from of a convolution with the Fourier-transformed weights function $W(\boldsymbol{k})$,
    \begin{equation}
        P^{W}_{\rm Tg}(k_i,\mu_i) = \frac{1}{V}\,\sum_j\,W(\boldsymbol{k}_i-\boldsymbol{k}_j)\,P_{\rm Tg}^{\rm obs\,eff}(k_j, \mu_j)\:.
    \end{equation}
    This model is used to generate a fiducial data vector. We have validated it against a direct method, where the input matter power spectrum is used to generate a log-normal, adding all effects on the level of the simulated vector. This allows us to propagate any statistical fields through the pipeline and estimate the power spectrum covariance from the sample covariance of the simulated spectra.
\section{Theory Model for BAO extraction}
    To extract cosmological information from the BAO, we introduce the two Alcock--Paczyński parameters $\alpha_{||}$ and $\alpha_\perp$, which are related to the expansion history of the universe. These parameters can be used to translate from "fiducial" cosmology to the "true" cosmology.
    \begin{align}
        k'(k, \mu) &= k\,\sqrt{\frac{1 - \mu^2}{\alpha^2_\perp} + \frac{\mu^2}{\alpha_{||}^2}}\:,\\
        \mu'(k, \mu) &=  \frac{\mu}{\alpha_{||}}\,\frac{1}{\sqrt{\frac{1 - \mu^2}{\alpha^2_\perp} + \frac{\mu^2}{\alpha_{||}^2}}}\:.
    \end{align}
    In the baseline model the theory is then fitted to the data shifting all scales:
    \begin{equation}
        P^{\rm th}(k_i, \mu_i) = P^{W}_{\rm gT}(k'(k_i, \mu_i), \mu'(k_i, \mu_i))\:.
    \end{equation}
    The dimensionality of the data is further reduced by binning the modulus of the wave vectors into one-dimensional bins. This is similar to extracting the cosmological monopole, but, due to the discretized wave vectors and non-cubic geometry, this quantity deviates from the true monopole through putting more weight into a preferred direction.
    \begin{equation}
        P^{\rm th}_{\rm 1d}(k_i) = \frac{1}{N_i}\,\sum_{\boldsymbol{k}_j \in \text{bin}_i} P^{\rm th}(k_j, \mu_j)
    \end{equation}
    To be more agnostic to possible systematics, which might change the overall shape of the power spectrum, we perform a marginalisation over the broad band shape of the power spectrum. This allows us to isolate the cosmological information in the position of the BAO.\\
    The broadband is given as a linear combination of multiple cubic splines.
    \begin{equation}
        X(x) = \frac{1}{6}\begin{cases}
            4 - 6\,|x|^2 + 3\,|x|^3& \text{, for }\:|x|<1\\
            8 - 12 \,|x| + 6\,|x|^2 - |x|^3& \text{, for }\:1\leq|x|<2\\
            0& \text{elsewise}
        \end{cases}
    \end{equation}
    By shifting the cubic spline we can achieve a smooth coverage of the range of scales.
    \begin{equation}
        \mathcal{D}(k_i) = \sum_{j=-1} a_i\,X\left(\frac{k_i}{\Delta}-j\right)\:,
    \end{equation}
    where the parameter $\Delta$ defines the width of the splines and thus the smallest feature that the broadband parametrisation would resolve. We fix $\Delta$ to a reasonable value larger than the distance between two BAO peaks. The amplitude parameters $a_i$ are free to vary and will be margenalised over.\\
    Having the freedom to margenalise over any smooth functions allows us to choose to model the power spectrum in a way that we explicitly remove any possible information from the broadband. In this model we write
    \begin{align}
        P^{\rm obs\,eff}_{\rm mod}(k_i, \mu_i) &= \mathcal{B}(k_i, \mu_i)\, \mathcal{P}(k_i, \mu_i)\, \mathcal{G}^2(k_i, \mu_i) \nonumber\\
        &\times \left[\mathrm{RSD}(k_i, \mu_i)\,P_{\rm nw}(k_i, \mu_i)\right.\nonumber\\
        &+\left.\mathrm{RSD}(k'_i, \mu'_i)\,g(k'_i, \mu'_i)\,P_{\rm w}(k'_i, \mu'_i)\right]\:,
    \end{align}
    where we have defined ${\rm RSD}$ as the combination of fingers-of-god effect and the two distinct kaiser terms. To obtain the model power spectrum, we need to convolve it with the survey weights and bin it like previously
    \begin{equation}
        P_{\rm 1d}^{\rm mod}(k_i) = \frac{1}{N_i}\sum_{\boldsymbol{k}_j \in \text{bin}_i}\,\frac{1}{V}\sum_{l}W(\boldsymbol{k}_j - \boldsymbol{k}_l)\,P^{\rm obs\, eff}_{\rm mod}(k_l, \mu_l)\:.
    \end{equation}
\section{Likelihood}
    To formulate our log-Likelihood is now given as the difference between the model power spectrum plus the broadband and the theoretical power spectrum which isolates the cosmological information coming only from the shifts in the apparent scale of the BAO wiggles. We write the Likelihood as
    \begin{align}
        \Delta_i &= P_{\rm 1d}^{\rm mod}(k_i) + \mathcal{D} - P^{\rm th}_{\rm 1d}(k_i),\\
        \log \mathcal{L} &= -\frac{1}{2}\, \Delta_i \mathrm{Cov}^{-1}_{ij} \, \Delta_j\:.
    \end{align}
    To speed up convergence and remove possible prior effects, we analytically margenalise over the $a_i$. Following standard procedure 
    \begin{align}
        \boldsymbol{X} &= \left[ X\left(\frac{k_i}{\Delta}+1\right), \dots, X\left(\frac{k_i}{\Delta}-N\right) \right]\:,\\
        \mathrm{Cov}^{-1}_{\rm marg} &=  \mathrm{Cov}^{-1} - \mathrm{Cov}^{-1}\,\boldsymbol{X}^\mathrm{T}\, \left(\boldsymbol{X}\,\mathrm{Cov}^{-1}\,\boldsymbol{X}^{\rm T}\right)^{-1}\,\boldsymbol{X}\,\mathrm{Cov}^{-1}
    \end{align}
    With this, we find a lower-dimensional expression for the Likelihood
    \begin{equation}
        \log \mathcal{L} = -\frac{1}{2}\left(P_{\rm 1d}^{\rm mod}- P^{\rm th}_{\rm 1d}\right) \mathrm{Cov}_\mathrm{marg}^{-1} \left(P_{\rm 1d}^{\rm mod}- P^{\rm th}_{\rm 1d}\right)^{\rm T}\:.
    \end{equation}
\end{document}
